### 2015-12-15 tex/draft2/ms.tex
update

%\documentclass[12pt, preprint,numberedappendix]{emulateapj}
%\documentclass[12pt, preprint]{aastex}
%\documentclass[dvips,12pt]{article}
\documentclass[apj]{emulateapj}

\newcommand\submitms{n}		% set to y to follow AAS ``ms'' names, etc.
\newcommand\bibinc{n}		% set to y if bib pasted in .tex, set to n to use bibtex

%\usepackage{pdfsync}
\usepackage{subeqnarray}
\usepackage{natbib}
\usepackage{color}
\usepackage[utf8]{inputenc}

\bibliographystyle{apj}

\newcommand{\ie}{i.e.\ }
\newcommand{\eg}{e.g.\ }
\newcommand{\p}{\partial}
\newcommand{\brak}[1]{\langle #1\rangle}

\newcommand{\gcc}{\;\mathrm{g\; cm^{-3}}}
\newcommand{\gsc}{\;\mathrm{g\; cm^{-2}}}
\newcommand{\cm}{\; {\rm cm}}
\newcommand{\mm}{\; {\rm mm}}
%\newcommand{\ps}{\; {\rm s^{-1}}}
\newcommand{\km}{\; {\rm km}}
\newcommand{\au}{\; \varpi_{\rm AU}}
\newcommand{\AU}{\; {\rm AU}}
\def\K{\; {\rm K}}

\newcommand{\vcs}[1]{\mbox{\boldmath{$\scriptstyle{#1}$}}}
\newcommand{\vc}[1]{\mbox{\boldmath{$#1$}}}
\newcommand{\nab}{\vc{\nabla}}
\DeclareMathSymbol{\varOmega}{\mathord}{letters}{"0A}
\DeclareMathSymbol{\varSigma}{\mathord}{letters}{"06}
\DeclareMathSymbol{\varPsi}{\mathord}{letters}{"09}

\newcommand{\Eq}[1]{Equation\,(\ref{#1})}
\newcommand{\Eqs}[2]{Equations (\ref{#1}) and~(\ref{#2})}
\newcommand{\Eqss}[2]{Equations (\ref{#1})--(\ref{#2})}
\newcommand{\App}[1]{Appendix~\ref{#1}}
\newcommand{\Sec}[1]{Sect.~\ref{#1}}
\newcommand{\Chap}[1]{Chapter~\ref{#1}}
\newcommand{\Fig}[1]{Fig.~\ref{#1}}
\newcommand{\Figs}[2]{Figs.~\ref{#1} and \ref{#2}}
\newcommand{\Figss}[2]{Figs.~\ref{#1}--\ref{#2}} 
\newcommand{\Tab}[1]{Table \ref{#1}}

\definecolor{gray}{gray}{0.5}
\newcommand{\emgr}[1]{\emph{ \color{gray} #1}}

%\newenvironment{packed_item}{
%\begin{itemize}
%  \setlength{\itemsep}{1pt}
%  \setlength{\parskip}{0pt}
%  \setlength{\parsep}{0pt}
%}{\end{itemize}}

\begin{document}

%\slugcomment{Draft Modified \today}

\title{The Role of Ice Compositions and Morphology for Snowlines and the C/N/O Ratios in Active Disks}

\author{Ana-Maria A. Piso\altaffilmark{1}, Karin I. \"Oberg\altaffilmark{1}, Jamila Pegues\altaffilmark{2}}
\altaffiltext{1}{Harvard-Smithsonian Center for Astrophysics, 60 Garden Street, Cambridge, MA 02138}
\altaffiltext{2}{Department of Astrophysical Sciences, Princeton University}

\begin{abstract}
The elemental compositions of planets define their chemistry, and could potentially be used as beacons for their formation location if the elemental gas and grain ratios of planet birth environments, i.e. protoplanetary disks, are well understood. In disks, the ratios of volatile elements, such as C/O and N/O, are regulated by the abundance of the main C, N, O carriers, the ice environment in which these carriers reside, and the presence of snowlines of major volatiles at different distances from the central star. We explore the effects of dynamical processes, molecular compositions and abundances, and the ice morphology of dust grains in disks on the snowline locations of the main C, O and N carriers, and their consequences for the C/N/O ratio in gas and dust throughout the disk. We find that radial drift and accretion alone can reduce the snowline radii of the main C, O and N carriers, i.e. H$_2$O, CO$_2$, CO and N$_2$, by 40-60\% compared to static disks. If CO and N$_2$ are bound to water ice instead of pure ices, their snowlines move inward by $\sim$$70$\%. Both of these effects substantially change the disk regions where C/O and N/O are enhanced over the stellar value. In the outer disk, the gaseous C/O and N/O are enhanced by factors of $\sim$$2$ and $\sim$$3$, respectively. Our estimates for the C/N/O ratios are only modestly affected by the presence of some C in the form of CH$_4$ and of some N in the form of NH$_3$.

%We note that N/O enhancements in disk gas can be even more extreme than C/O in the outer disk due to the low volatility of N2 compared to all major C and O carriers. I will discuss these results together with the effects of additional dynamical processes, and outline a path toward a coupled drift-desorption-chemistry model that will provide robust quantitative results for volatile snowline locations and C/N/O abundance ratios as the disk evolves in time.
\end{abstract}

\section{Introduction}
\label{sec:intro}

\emgr{Background info. Importance of volatiles in disks and planetary atmospheres, detections of snowlines in disks, C/O ratios etc. State again the importance of radial drift and gas accretion on the snowline locations, and that a systematic study of the combination of these two particular effects across the disk has not been done before. Then transition to the fact that we provide such a systematic study in Paper I and in this paper. Here, we expand the model of Paper I by making three additions: (1) we add N and CH4 in the static chemistry model, and explore how different abundances of CH4 and of the N main carriers (N2 and NH3) affect the C/O  and N/O ratios, (2) we quantify the effect of radial drift and gas accretion on the N2, CH4 and NH3 snowline locations, and (3) we explore how different binding energies of CO and N2 affect their snowline locations.}

\begin{figure*}[t!]
\centering
\includegraphics[width=\textwidth]{../../figs/CNO_and_snowlines.pdf}
%\vspace{-0.5in}
\caption{CNO abundance in grains...} 
\label{fig:CNOstatic}
\end{figure*}

\section{Coupled Drift-Desorption Model Review}
\label{sec:review}

We begin with a brief review of Paper I's model for the effect of radial drift and viscous gas accretion on volatile snowline locations. We review our disk model in Section \ref{sec:disktime}, and summarize our numerical method and results in Section \ref{sec:driftdes}.

\subsection{Disk Model}
\label{sec:disktime}

We first assume a static disk, which is only irradiated by the central star and does not experience redistribution of solids or radial movement of the nebular gas. To quantify the effects of radial drift and gas accretion, we use a viscous disk with a spatially and temporally constant mass flux, $\dot{M}$. The viscous disk takes into account radial drift, gas accretion onto the central star, as well as accretion heating. We prefer this disk model to an irradiated or evolving disk (see Paper I) because it includes all the dynamical and thermal processes we are interested in for the scope of this paper, and therefore it is the most realistic one.

We model the static disk as a minimum mass solar nebula (MMSN), using a prescription for the gas surface density, $\Sigma$, and disk midplane temperature, $T$, similar to that of \citet{chiang10}:
\begin{subeqnarray}
\label{eq:disk}
\Sigma&=&2000\, (r/\text{AU})^{-1}\,\, \text{g cm}^{-2} \slabel{eq:disksigma}\\
T &=& 120\, (r/\text{AU})^{-3/7} \,\,\text{K}, \slabel{eq:diskT}
\end{subeqnarray}
where $r$ is the semimajor axis. Based on observations of protoplanetary disks \citep{andrews10}, we choose a flatter surface density than the one assumed by \citet{chiang10}, where $\Sigma \propto r^{-3/2}$.

We use the \citet{shakura73}  steady-state disk solution to model the viscous disk. Solving the Equation set of Appendix A in Paper I yields an expression for the temperature profile in a steady-state disk:
\begin{equation}
\label{eq:Tdact}
T_{\rm act}=\frac{1}{4 r} \Big(\frac{3 G \kappa_0\dot{M}^2 M_* \mu m_{\rm p} \Omega_{\rm k}}{\pi^2 \alpha k_{\rm B} \sigma}\Big)^{1/3}.
\end{equation}
Here $G$ is the gravitational constant, $\kappa_0=2 \times 10^{-6}$ is a dimensionless coefficient in the opacity law $\kappa=\kappa_0 T_{\rm act}^2$, $M_*=M_{\odot}$ is the mass of the central star, $\mu=2.35$ is the mean molecular weight of the nebular gas, $m_{\rm p}$ is the proton mass, $\Omega_{\rm k}=\sqrt{G M_{\odot}/r^3}$ is the Keplerian angular velocity, $\alpha=0.01$ is a dimensionless coefficient (see below for details), $k_{\rm B}$ is the Boltzmann constant, and $\sigma$ is the Stefan-Boltzmann constant. The final midplane temperature profile is computed as 
\begin{equation}
\label{eq:activeT}
T^4 = T_{\rm act}^4 + T_{\rm irr}^4,
\end{equation}
where $T_{\rm irr}=T$ from Equation (\ref{eq:diskT}). We use this expression because in addition to accretion heating, stellar irradiation also contributes to the disk thermal structure.

The steady-state disk has an $\alpha$-viscosity prescription, where the kinematic viscosity is $\mu=\alpha c H$. Here $c \equiv \sqrt{k_{\rm B} T /(\mu m_{\rm p})}$ is the isothermal sound speed (with $T$ from Equation \ref{eq:activeT}), and $H \equiv c/\Omega_{\rm k}$ is the disk scale height. We can then determine the gas surface density for a viscous disk as (\citealt{shakura73}; see also Paper I for a more detailed explanation of these calculations):
\begin{equation}
\label{eq:Sigmaact}
\Sigma=\frac{\dot{M}}{3 \pi \nu}.
\end{equation}
We choose $\dot{M}=10^{-8} M_{\odot}$ yr$^{-1}$, consistent with mass flux observations in disks (e.g., \citealt{andrews10}). As acknowledged in Paper I, the mass flux rate $\dot{M}$ and stellar luminosity $L_*$ will vary throughout the disk lifetime (\citealt{kennedy06}, \citealt{chambers09}), in contrast with our simplified model which assumes that both quantities are constant. This effect will be most pronounced in the inner disk ($\lesssim$ few AU), where accretion heating dominates. We thus acknowledge that the location of the H$_2$O snowline may be determined by the decline in $\dot{M}$ or $L_*$ with time, rather than radial drift (see Paper I, Section 2.1 for a more detailed explanation).

%In what follows, we summarize the relevant timescales that affect volatile snowline locations. 
%
%\textit{Desorption timescale.} Following \citet{hollenbach09} and Paper I, the timescale to desorb a single layer of molecules of a volatile $x$ can be approximated as
%\begin{equation}
%\label{eq:tdes}
%t_{\rm des}=\frac{\rho_{\rm s}}{3 \mu_x m_{\rm p}} \frac{s}{N_x R_{\rm des, x}},
%\end{equation} 
%where $\rho_{\rm s}=2$ g cm$^{-3}$ is the density of an icy particle, $s$ is the particle size, $\mu_x$ is the mean molecular weight of volatile $x$, and $N_x \approx 10^{15}$ sites cm $^{-2}$ is the number of adsorption sites of molecule $x$ per cm$^{-2}$. The desorption rate $R_{\rm des, x}$ (per molecule) is \citep{hollenbach09}
%\begin{equation}
%\label{eq:Rdes}
%R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
%\end{equation}
%where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}=T$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well. We discuss our choices for $E_x$ for the different volatile species in Sections \ref{sec:CH4} and \ref{sec:N}.

\subsection{Desorption-Drift Equations and Results}
\label{sec:driftdes}

For a range of initial icy grain sizes composed of a single volatile, we showed in Paper I that the timescale on which these particles desorb is comparable to their radial drift time, as well as to the accretion timescale of the nebular gas onto the central star. We thus have to take into account both drift and gas accretion when we calculate the disk location at which a particle desorbs, since that location may be different from the snowline position in a static disk for a given volatile (see Figure \ref{fig:CNOstatic} and \citealt{oberg11}). We determine a particle's final location in the disk by solving the following coupled differential equations:
\begin{subeqnarray}
\label{eq:ddt}
\frac{ds}{dt} &= & - \frac{3 \mu_x m_{\rm p}}{\rho_{\rm s}} N_x R_{\rm des, x}  \slabel{eq:dsdt} \\
\frac{dr}{dt} &=& \dot{r} \slabel{eq:drdt},
\end{subeqnarray}
where $s$ is the particle size, $t$ is time, $\mu_x$ is the mean molecular weight of volatile $x$, $\rho_{\rm s}=2$ g cm $^{-3}$ is the density of an icy particle, $N_x \approx 10^{15}$ sites cm $^{-2}$ is the number of adsorption sites of molecule $x$ per cm$^{-2}$, $R_{\rm des, x}$ is the desorption rate of species $x$, and $\dot{r}$ is the particle's radial drift velocity. We calculate $R_{\rm, des}$ and $\dot{r}$ as follows.

The desorption rate $R_{\rm des, x}$ (per molecule) is \citep{hollenbach09}
\begin{equation}
\label{eq:Rdes}
R_{\rm{des}, x} = \nu_x \exp{(-E_x/T_{\rm grain})},
\end{equation}
where $E_x$ is the adsorption binding energy in units of Kelvin, $T_{\rm grain}=T$ is the grain temperature (assumed to be the same as the disk temperature, see Paper I), and $\nu_x=1.6 \times 10^{11} \sqrt{(E_x/\mu_x)}$ s$^{-1}$ is the molecule's vibrational frequency in the surface potential well. We discuss our choices for $E_x$ for the different volatile species in Sections \ref{sec:CH4} and \ref{sec:N}.

Following \citet{chiang10} and \citet{birnstiel12}, a particle's radial drift velocity can be approximated as 
\begin{equation}
\label{eq:rdotact}
\dot{r} \approx -2 \eta \Omega_{\rm k} r \Big(\frac{\tau_{\rm s}}{1+\tau_{\rm s}^2}\Big) + \frac{\dot{r}_{\rm gas}}{1+\tau_{\rm s}^2},
\end{equation}
where the first term is the drift velocity in a non-accreting disk and the second term accounts for the radial movement of the gas. Here $\eta \approx c^2/(2 v_{\rm k}^2)$, where $v_{\rm k}$ is the Keplerian velocity, and $\tau_{\rm s} \equiv \Omega_{\rm k} t_{\rm s}$ is the dimensionless stopping time:
\begin{equation}
\label{eq:ts}
t_{\rm s}= \left\{
\begin{array}{l l}
\rho_{\rm s} s / (\rho c), & \quad s < 9 \lambda/4 \,\,\,\ \text{Epstein drag} \\
4 \rho_{\rm s} s^2 / (9 \rho c \lambda), & \quad s < 9 \lambda/4, \,\text{Re} \lesssim 1 \,\,\,\ \text{Stokes drag,}
\end{array} 
\right.
\end{equation}
where $\rho$ is the disk mid-plane density, $\lambda$ is the mean free path and Re is the Reynolds number.  The gas accretion velocity $\dot{r}_{\rm gas}$ is determined from $\dot{M}=-2 \pi r \dot{r}_{\rm gas} \Sigma$, for a fixed $\dot{M}$ and with $\Sigma$ given by Equation (\ref{eq:Sigmaact}).

For a particle of initial size $s_0$, we solve the Equation set (\ref{eq:ddt}) with the initial conditions $s(t_0)=s_0$ and $r(t_0)=r_0$, where $t_0$ is the time at which we start the integration and $r_0$ is the particle's initial location. We stop our simulation after $t_{\rm d}=3$ Myr, the disk lifetime, since this is roughly the timescale on which planets form \citep{}, and determine the desorption timescale $t_{\rm des}$ from $s(t_{\rm des})=0$, and thus a particle's desorption distance $r_{\rm des}=r(t_{\rm des})$. Our results are insensitive to our choice of $t_0$ as long as $t_0 \ll t_{\rm d}$. We note that a particle's size is initially fixed and only changes due to desorption. We thus do not take into account processes such as grain coagulation or fragmentation, which nonetheless occur in disks (e.g., \citealt{birnstiel12}, \citealt{perez12}). We discuss the effect of these processes on snowline locations in Paper I.

As we show in Paper I, a particle of initial size $s_0$ can experience three outcomes after $t_{\rm d}=3$ Myr: (1) it can remain at its initial location, (2) it can drift towards the host star, then stop without evaporating significantly, and (3) it can completely desorb on a timescale shorter than 3 Myr. Particles in scenarios (1) and (2) are thus not affected by radial drift or gas accretion, and the snowline locations are those for a static disk. In contrast, the grains in case (3) desorb \textit{instantaneously} and \textit{at a fixed particle-size dependent location} in the disk, regardless of their initial position. The snowline locations for these particles will thus be fixed for a given initial particle size and disk model. We have found that grains with sizes $\sim0.001$ cm $\lesssim s \lesssim$ 7 satisfy this condition for our fiducial disk.

%\emgr{Review disk models, desorption model, relevant timescales. State that we use a steady-state disk for the coupled drift-desorption evolution, since it is the most realistic, therefore only summarize the static and steady-state (viscous) disk. Summarize the findings of Paper I, i.e. particles of certain sizes desorb instantaneously and at a fixed particle size dependent location.}

\section{CH$_4$ and C/O Ratios}
\label{sec:CH4}

Both in Solar system comets and in protoplanetary disks, carbon and oxygen are primarily contained in H$_2$O, CO$_2$ and CO (e.g., \citealt{rodgers02}, \citealt{lodders03}, \citealt{pontoppidan06}). However, some fraction of the carbon abundance may also be carried by CH$_4$ (e.g., \citealt{mumma96}), which may change the C/O ratio in gas and in dust throughout the disk. To quantify the magnitude of this effect, we use measured CH$_4$ abundances in protostellar cores from the \textit{Spitzer} c2d Legacy ice survey \citep{evans03}. We explore the parameter space of possible CH$_4$ abundances by assuming three different scenarios: (1) no CH$_4$, (2) the median CH$_4$ observed abundance (hereafter CH$_4$-mid), and (3) the maximum CH$_4$ observed abundance (hereafter CH$_4$-max). Thus $n_{\rm CH_4-mid}=0.0555 \times n_{\rm H_2O}$ \citep{oberg11a} and $n_{\rm CH_4-mid}=0.13 \times n_{\rm H_2O}$ \citep{oberg08}. Similarly to Paper I, we use the H$_2$O, CO$_2$ and CO abundances of \citet{oberg11}. Since the abundance of carbon grains is uncertain, we assume that all the carbon that is not in the form of CH$_4$ is found in carbon grains, so that we reproduce the Solar C/O ratio (gas+dust) of 0.54.

We determine the location of the H$_2$O, CO$_2$, CO and CH$_4$ snowlines in our static disk by balancing desorption with readsorption, following \citet{hollenbach09}. The binding energies of H$_2$O, CO$_2$, CO and CH$_4$ as pure ices are 5800 K, 2000 K, 1388 K and 1300 K, respectively (\citealt{fraser01}, \citealt{collings04}, Fayolle et al., submittted, \citealt{garrod06}).

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/C_O_ratio_CH4.pdf}
%\vspace{-0.5in}
\caption{C/O ratio in a static disk for different CH4 abundances and CO binding energies...} 
\label{fig:COstatic}
\end{figure}

Figure \ref{fig:COstatic} shows the C/O ratio in gas and dust as a function of semimajor axis in a static disk, for different CH$_4$ abundances as outlined above. As in \citet{oberg11} and Paper I, a gaseous C/O ratio of unity can be achieved beyond the CO$_2$ snowline, where oxygen gas is significantly depleted (top panel). The gas-phase C/O ratio may be further enhanced between the CO$_2$ and CH$_4$ snowlines due to the presence of additional carbon gas from CH$_4$. In this region, the C/O ratio increases by 3\% for CH$_4$-mid and by 8\% for  CH$_4$-max, as displayed in the middle and bottom panel of Figure \ref{fig:COstatic}. Based on current observations of CH$_4$ abundances, its presence in the disk only modestly affects the C/O ratio. Given the larger uncertainties in overall volatile abundances, we can neglect CH$_4$ when estimating the C/O ratio in static disks %\footnote{This may not be the case, however, in a viscous disk, as we show below.}.

As noted in Section \ref{sec:intro}, the CO binding energy varies significantly depending on the environment in which the icy grains reside. If CO ice is layered on top of a water ice substrate, its binding energy will be larger than in the pure ice case (1388 K) due to the higher H$_2$O binding energy. Fayolle et al. (submitted) find a CO binding energy of 834 K in the layered ice scenario. We use the model of Section \ref{sec:review} to estimate the movement of the CO snowline for different grain morphologies in a viscous disk.

Figure \ref{fig:CO_ratio} shows the H$_2$O, CO$_2$ and CO snowline locations for particles with initial sizes $\sim0.06$ cm $\lesssim s \lesssim$ 7 m as well as estimates for the C/O ratio in gas and dust in a viscous disk, with the CO snowline calculated under different grain morphologies as noted above. The true snowline for particles that desorb outside the static snowline is the static snowline itself, hence desorbing particles with $s<0.06$ cm do not form true snowlines. As calculated in Paper I, drift and gas accretion may move the snowlines inwards by up to 40-60\% compared to a static disk, and specifically by up to $\sim50$ \% in the case of the CO snowline. This result is preserved for the updated CO binding energies, both for pure ice and on a water ice substrate. However, a layered grain structure moves the CO snowline inwards significantly: for our fiducial disk model, $r_{\rm CO, layered} \approx 8.7$ AU. Thus if CO is bound to water ice instead of pure ices, the CO snowline may move inward by up to 70\%. This large inward movement of the CO snowline implies that C/O ratios of order unity may be reached much closer to the host star if CO is layered on a water ice substrate, and may be inside 10 AU for certain disk parameters.

%\emgr{Discuss observed abundances for CH4 and the choices that we make (no CH4, median value, maximum value). State that desorption energies for H2O, CO2 and CH4 are well constrained experimentally, and that the CO2 and CH4 binding energies are only weakly dependent on whether it's pure CO2/CH4 or combined with H2O, but that is not the case for CO (and N2 as we will show in the next section). Present new binding energies for CO as pure ice and mixed with water. Show Figure 1 and discuss how different CH4 abundances and binding energies affect snowline locations and C/O ratio: CO-H2O mixture (though I think it's rather CO layered on top of H2O) moves the CO snowline inward by $\sim$40 AU (will calculate percentages too); the maximum reasonable abundance of CH4 changes the C/O ratio by less than 10\%. Show Figure 2 and quantify the effect of drift and accretion on the CH4 snowline compared to a static disk. While CH4 has only a modest effect on the C/O ratio in a static disk, this effect may be larger in a viscous disk, as the C gas abundances inside the CH4 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I). In this study, however, we neglect these effects and therefore do not include CH4 in estimating the C/O ratio (as an aside, the figures that include CH4 in the C/O ratio with drift and desorption are quite messy due to snowlines overlapping). Show Figure 3 and estimate the difference between CO-H2O and CO pure ice snowlines in the case of drift and accretion, as well as the comparisons for the static disk for the CO snowline.}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/C_O_water_ice.pdf}
%\vspace{-0.5in}
\caption{C/O ratio as function of semimajor axis for CO combined with H2O (top panel) and pure CO ice (bottom panel).... Drift and gas accretion move the CO snowlines inward by x\% and y\%, respectively.} 
\label{fig:CO_ratio}
\end{figure}

\section{Nitrogen Carriers and N/O Ratios}
\label{sec:N}

\emgr{Similar to the previous section, but with more details. Discuss that nitrogen is abundant in the solar system and disks and primarily found as N2. Due to the high volatility of N2, the gas phase N/O ratio in the outer disk may be even more enhanced than the C/O ratio. A fraction of the nitrogen abundance may be also carried by NH3. Discuss NH3 observed abundances and the choices that we make (no NH3, median, maximum). State that the NH3 desorption energy is only weakly dependent on whether it's pure NH3 or combined with H2O, but that is not the case for N2. Present new binding energies for N2 as pure ice and combined with water. Show Figure 4 and discuss how different nitrogen abundances and binding energies affect snowline locations and N/O ratio: N2 combined with H2O moves the N2 snowline inward by $\sim$50 AU (will calculate percentages too); the maximum reasonable abundance of NH3 changes the N/O ratio by $\sim$15\%. In the outer disk, the N/O ratio is enhanced by a factor of $\sim$4 compared to the solar value, twice as much as the C/O enhancement. Show Figure 5 and quantify the effect of drift and accretion on the NH3 snowline compared to a static disk. While NH3 does not have a significant effect on the N/O ratio in a static disk, this effect may be larger in a viscous disk, as the N gas abundance inside the NH3 snowline may be enhanced due to the differential motion of the desorbed ices and overall nebular gas (refer to Paper I). In this study, however, we neglect these effects and therefore do not include NH3 in estimating the N/O ratio (again, N/O ratio figure with drift is quite messy when including NH3 and does not add any information that is not already shown in Figures 4 and 5). Show Figure 6 and estimate the difference between N2-H2O and N2 pure ice snowlines in the case of drift, as well as the comparisons for the static disk for the N2 snowline. State that there will be an overabundance of gas-phase N/O between the CO and N2 snowlines, as there is no oxygen gas in this region.}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/N_O_ratio.pdf}
%\vspace{-0.5in}
\caption{N/O ratio in a static disk for different NH3 abundances and N2 binding energies...} 
\label{fig:Nstatic}
\end{figure}

%\begin{figure}[h!]
%\centering
%\includegraphics[width=0.5\textwidth]{../../figs/desorption_distance_NH3_steady.pdf}
%%\vspace{-0.5in}
%\caption{Desorption distance as a function of initial distance for NH3.... Drift and gas accretion move the NH3 snowline inward by x\%.} 
%\label{fig:NH3}
%\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=0.5\textwidth]{../../figs/N_O_water_ice.pdf}
%\vspace{-0.5in}
\caption{N/O ratio as function of semimajor axis for N2 combined with H2O (top panel) and pure N2 ice (bottom panel).... Drift and gas accretion move the N2 snowlines inward by x\% and y\%, respectively. Overabundance of gas-phase N/O between the CO and N2 snowlines, marked by the vertical red dash-dotted lines for the largest drifting particles in our model.} 
\label{fig:NO_ratio}
\end{figure}

\section{Discussion}
\label{sec:discussion}

\emgr{Discuss how entrapment of volatiles by H2O affects volatile abundances and C/O ratios. Re-emphasize the fact that the C/O and N/O ratios are upper estimates, and that CH4 and NH3 might matter in a viscous disk. State that we plan to address this in a future paper. More TBD.}

\section{Summary}

\emgr{Maybe we can include the summary in the discussion section?}

%\section{Volatile Abundances and Binding Energies: Effect on C/O and N/O Ratios in a Static Disk}
%
%\subsection{Nitrogen and CH4}
%
%\emgr{Discuss that nitrogen is abundant in the solar system and should be abundant in disks as well, but that its dominant form is largely unknown. Discuss that the main carrier of nitrogen is N2, but some fraction of it can be NH3 as well. Present and motivate the choices that we make for NH3 abundances. Along the same lines, discuss CH4, and the choices that we make for CH4 abundances.}
%
%\subsection{Volatile Desorption Energies}
%
%\emgr{State that desorption energies for H2O and CO2 are well constrained experimentally, and that the CO2 binding energy is only weakly dependent on whether it's pure CO2 or combined with H2O, but that is not the case for CO, N2 (and perhaps CH4 and NH3?). Briefly discuss CO-CO, N2-N2, CO-H2O and N2-H2O (and perhaps the same for CH4 and NH3 if we find literature on that) experimental results for binding energies, and motivate the choices that we make.}
%
%\subsection{Results for C/O and N/O in a Static Disk}
%
%\emgr{For each of them (C/O and N/O), show a 3-panel plot as follows: each panel has a specific CH4/NH3 abundance (top: none, middle: median abundance, bottom: maximum abundance); for a given panel, have multiple curves for C/O or N/O, depending on the choice of binding energies, so that it's clear visually how the binding energy changes the snowline location. Discuss how different abundances  and binding energies affect snowline locations and C/O or N/O ratios.}
%\section{Results}
%
%\subsection{N2, NH3 and CH4 Snowline Locations}
%
%\emgr{One multipanel 3x3 (or 3x2) rainbow plot similar to the snowline plots from Paper I, for \textit{one} choice for the binding energies (perhaps the largest ones, since we want a limit on how far in we can push the snowlines?). Rows: snowlines as a function or particle size for passive, active and (maybe) steady-state disk. Columns: N2, NH3 and CH4. Not entirely certain how necessary these plots/subsection actually are, since it is exactly the same qualitatively as in Paper I...}
%
%\subsection{C/O and N/O Ratios}
%
%\emgr{For each of them (C/O and N/O) show a 3x3 multipanel plot similar to the C/O plot from Paper I, for the same choice of binding energies as in the previous subsection. Columns: C/O or N/O in passive, active and steady-state disk. Rows: No CH4/NH3, median CH4/NH3, maximum CH4/NH3. Again, not sure if we need/want this for all disk choices, at least in the case of C/O since we already have that in Paper I. For N/O, quantify how the snowline location changes from a static disk due to drift and gas accretion.}

\if\bibinc n
\bibliography{refs}
\fi

\if\bibinc y
\begin{thebibliography}
\end{thebibliography}
\fi

\end{document}