### 2016-04-15 Dissertate-Harvard-LaTeX/frontmatter/thanks.tex

% the acknowledgments section

First and foremost, I would like to thank Karin \"Oberg for all the support she has provided me with in the past three years, one as my mentor and the last two as my thesis advisor. None of this would have been possible without her constant guidance and encouragement. I cannot think of a moment when she hasn't been responsive, be it to bounce off ideas, to give me feedback on a paper, to counsel me regarding professional advancement, or to offer incredible support and understanding during the tougher times. I've learnt from her not only how to be a successful scientist, but also what qualities make one an exceptional advisor and mentor. I am grateful that Karin joined the Harvard faculty three years ago; it's been a privilege to work with her.

Secondly, I would like to offer my thanks to Ruth Murray-Clay for advising me during the first few years of my PhD. She guided me through my first steps as a professional astronomer, and she always encouraged me to become more independent as a scientist and to not be afraid to speak my mind. When I get to teach my first astronomy class, I will pleasantly remember the many hours Ruth and I spent doing order of magnitude calculations in her office or at group meetings.

I'd like to give a special thanks to Rosanne Di Stefano, my faculty mentor for the past two years. I've immensely enjoyed our weekly chats over breakfast at Sarah's or Hi-Rise. I am glad I could benefit from both her wisdom and her kindness when I needed either professional or personal advice. Even in the darkest moments, Rosanne has managed to put a smile on my face and truly make me believe that things always get better.

There are many, many other people at the CfA whom I'd like to thank and who made my graduate experience both productive and enjoyable. In particular, I would like to thank Sean Andrews, Matt Holman and Dimitar Sasselov for accepting to be part of my Thesis Advising Committee and for guiding me step by step towards a successful thesis completion. Although outside the CfA, I would like to thank Fred Ciesla for flying over from Chicago to be part of my Defense Committee, and for sharing excellent research ideas when I visited the University of Chicago last fall. I've also appreciated the occasional but insightful conversations with Dave Charbonneau, John Johnson and Alicia Soderberg. A special thanks goes to Til Birnstiel, with whom I've enjoyed collaborating and who would be almost instantly available when I had a question or I was stuck debugging my code. I would also like to thank the astrochemistry and disk groups for often offering me helpful suggestions and ideas, and for teaching me a thing or two about chemistry and radio astronomy.

My life in the Astronomy department has been better due to my fellow graduate students. Meredith MacGregor and Courtney Dressing have been wonderful office mates. I have often been inspired by their intelligence and hard working abilities, as much as I've enjoyed going to a movie or an after work drink with them. Anjali Tripathi has been an immense support throughout my time here; I cannot count the late nights we spent in one of our offices trying to help each other with some work problem or just talking about life.

This section would not be complete if I didn't thank my friends outside of astronomy, who have been there for me willing to give me a hug in the good times and especially during the bad ones. My (more frequent than I would like to admit) gchats with Andra about everything and nothing always make my day. The weekly Saturday morning breakfast with Carlo and the evenings of Walking Dead marathons have been the best, as were the frequent froyo runs with Luly. I will miss checking out new restaurants with Laura and Iulia, as I will miss John's smile and contagious optimism.

Finally, but perhaps most importantly, I would like to thank my family for loving and supporting me throughout these last ten years in Boston, even though they are so far away. My dad was the first to trigger my interest in astronomy by teaching me the names of the planets in our Solar System when I was around five. I don't have enough words to thank my mom. She is the most amazing person that I know; I would not have gotten anywhere close to here without her.